---
title: "Superstore-with-R"
author: "Yu Yuen Hern"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tableau Superstore Dataset

With growing demands and cut-throat competitions in the market, a Superstore Giant is seeking your knowledge in understanding what works best for them. They would like to understand which products, regions, categories and customer segments they should target or avoid.

They also want to have a Regression model to predict Sales or Profit.

# Get the dataset

Firstly, let's get the dataset from XLS file. In the XLS file, there are 3 sheets:

-   Orders: List of transactions
-   Returns: List of items returned
-   People: List of sales person for West, East, Central and South

Let's import the necessary libraries first.

```{r}
library(readxl)
```

Now let's retrieve the dataset in as a dataframe.

```{r, echo=FALSE}
df_order = read_xls('superstore.xls', sheet = 'Orders', col_names = TRUE)
df_return = read_xls('superstore.xls', sheet='Returns', col_names = TRUE)
df_people = read_xls('superstore.xls', sheet='People', col_names = TRUE)
```

## Checking the data



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
